\normalfont\spacedallcaps{Language reclamation: Literature review}

### 2023-06-19 assignment-4/article.tex
Completed solution to 16.2
Add a solution to 16.11

\begin{align*}
  p_X(x) &= p_Z(z) \left|\frac{dz}{dx}\right| \\
  &= \frac{1}{\sqrt{2\pi}} \exp \left[-\frac{1}{2}(\log(x(1-x)))^2 \right] \cdot \frac{1-2x}{x(1-x)} \\
  &= \frac{1}{\sqrt{2\pi}} \exp \left[\log(x(1-x)))^{-\frac{1}{2} \cdot 2} \right] \cdot \frac{1-2x}{x(1-x)} \\
  &= \frac{1}{\sqrt{2\pi}} \exp \left[\log\left(\frac{1}{x(1-x)}\right) \right] \cdot \frac{1-2x}{x(1-x)} \\
  &= \frac{1}{\sqrt{2\pi}} \frac{1}{x(1-x)} \cdot \frac{1-2x}{x(1-x)} \\
  &= \frac{1}{\sqrt{2\pi}} \frac{1-2x}{x^2(1-x)^2}
\end{align*}

%% \subsubsection*{Problem 16.2 solution - attempt 2}

%% Given the standard normal distribution:
%% \begin{equation*}
%%   Pr(z) = \frac{1}{\sqrt{2\pi}} \exp \left[-\frac{z^2}{2}\right]
%% \end{equation*}

%% And the function:
%% \begin{equation*}
%%   x = f[z] = \frac{1}{1 + \exp[-z]}
%% \end{equation*}

%% We can derive the inverse function $f^{-1}(x)$:
%% \begin{equation*}
%%   z = f^{-1}(x) = -\log \left(\frac{1}{x} - 1\right)
%% \end{equation*}

%% The formula for the change of variables in probability densities is:

%% \begin{equation*}
%%   Pr(x) = Pr(f^{-1}(x)) \left| \frac{df^{-1}(x)}{dx} \right|
%% \end{equation*}

%% By differentiating $f^{-1}(x)$ with respect to $x$, we get $\frac{df^{-1}(x)}{dx} = \frac{1}{x(1-x)}$.

%% Substituting $f^{-1}(x)$ and $\frac{df^{-1}(x)}{dx}$ into the formula, we get:

%% \begin{equation*}
%%   Pr(x) = \frac{1}{\sqrt{2\pi}} \exp \left[-\frac{(-\ln(\frac{1}{x} - 1))^2}{2}\right] \cdot \frac{1}{x(1-x)}
%% \end{equation*}

%% This is the probability density function of the transformed distribution.

@@

>>>>

\begin{align*}
  p_X(x) &= p_Z(z) \left|\frac{dz}{dx}\right| \\
  &= \frac{1}{\sqrt{2\pi}} \exp \left[-\frac{1}{2}(\log(x(1-x)))^2 \right] \cdot \frac{|1-2x|}{x(1-x)}
\end{align*}

The expression looks a bit nasty but I don't think it's possible to simplify further.

@@

\subsubsection*{Problem 16.11 solution}

Writing out the KL divergence in full gives:

\begin{equation*}
  \phi = \argmin \left[ \int \sum_{i=1}^I \delta[x-f[z_i,\phi]] \log \frac{\sum_{i=1}^I \delta[x-f[z_i,\phi]]}{q(x)} dx \right]
\end{equation*}

If we replace $q(x)$ with $\kappa q(x)$ in the above equation it can be factored out as an additive constant in the following way:

\begin{align*}
  \phi &= \argmin \left[ \int \sum_{i=1}^I \delta[x-f[z_i,\phi]] \log \frac{\sum_{i=1}^I \delta[x-f[z_i,\phi]]}{\kappa q(x)} dx \right] \\
  &= \argmin \left[ \int \sum_{i=1}^I \delta[x-f[z_i,\phi]] \log \frac{\sum_{i=1}^I \delta[x-f[z_i,\phi]]}{q(x)} dx - \log \kappa \int \sum_{i=1}^I \delta[x-f[z_i,\phi]] dx \right] \\
  &= \argmin \left[ \int \sum_{i=1}^I \delta[x-f[z_i,\phi]] \log \frac{\sum_{i=1}^I \delta[x-f[z_i,\phi]]}{q(x)} dx - \log \kappa \right]
\end{align*}

Where the last line holds since the probability density $\sum_{i-1}^I \delta[x-f[z_i,\phi]]$ integrates to 1 over its domain. Since the result of multiplying $q(x)$ by a constant $\kappa$ only adds a constant term to the KL-divergence, the $\phi$ which minimizes this divergence is the same.

The network does not need to be invertible to minimize the KL divergence. Since we generate the samples $x_i = f[z_i , \phi]$ directly from the latent variables $z_i$, we do not need to perform an inverse operation from the data space to the latent space, which would require the network to be invertible.
